### 2014-12-10 Rmarkdown/wgbs_statistical_framework.Rmd
Typos and latex errors
Updated appendix
Latex fixes

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/bs-seq_protocol.pdf}
\caption{A schematic illustrating several sources of variation in a bisulfite-sequencing experiment. Each sample starts as a population of cells, with potentially different methylation patterns. Each DNA fragment is coloured by its originating cell, although this is unknown in practice. Illustrated are PCR amplication bias (unequal representation of DNA from each cell following PCR amplification), sampling variation (not all DNA fragments are sequenced), alignment errors (not all sequenced fragments are mapped or may be mapped to the wrong location, as is the case for the green fragment), and filtering during methylation calling (not reads are used, the blue read, and reads may have read-positions removed, the black reads).}
\label{fig:bs-seq_protocol}
\end{figure}

@@

\begin{equation*}
Z_{h, i} = \left\{
  \begin{array}{l l}
  (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus on the } h^{th} \text{ fragment}\\
  (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th}) \text{ locus and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus on the } h^{th} \text{ fragment} \\
  \vdots \\
  (1, 1, \ldots, 1) & \quad \text{if methylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus on the } h^{th} \text{ fragment}
  \end{array} \right.
  \end{equation*}

@@

\begin{equation*}
z: z \in \mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)} = \left\{
  \begin{array}{l l}
    (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus}\\
    (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th}) \text{ locus and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus} \\
    \vdots \\
    (1, 1, \ldots, 1) & \quad \text{if methylated at the } i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th}) \text{ locus}
      \end{array} \right.
\end{equation*}

@@

- TODO: Lots of figures and discussion. See chapter plan.

@@

Restricting our attention to CpGs, Figures \ref{fig:EPISCOPE_plotBetaDensity, fig:Lister_plotBetaDensity, fig:Seisenberger_plotBetaDensity, fig:Zilelr_plotBetaDensity} show the kernel density estimates of the genome-wide distributions of $\beta$-values, that is, the column-wise summaries, for each sample of the EPISCOPE, Lister, Seisenberer and Ziller datasets, respectively. Figures \ref{fig:EPISCOPE_plotBetaFreqPoly, fig:Lister_plotBetaFreqPoly, fig:Seisenberger_plotBetaFreqPoly, fig:Zilelr_plotBetaFreqPoly} show the same data but with the $\beta$-values grouped into $0.01$-width bins and plotted against the percentages of CpGs that fall into each bin.

@@

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity}
\end{figure}

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Lister data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity}
\end{figure}

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity}
\end{figure}

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/Ziller_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Ziller data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_plotBetaDensity}
\end{figure}

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:EPISCOPE_plotBetaFreqPoly}
\end{figure}

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Lister data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:Lister_plotBetaFreqPoly}
\end{figure}

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:Seisenberger_plotBetaFreqPoly}
\end{figure}

\begin{figure}[htbp]
\includegraphics[width=\textwidth]{../figures/Ziller_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Ziller data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:Ziller_plotBetaFreqPoly}
\end{figure}

The bimodality of the genome-wide $\beta$-value distributions is driven by the fact that most CpGs is CpG islands are unmethylated whereas those outside of the islands are mostly methylated. Figures \ref{fig:EPISCOPE_plotBetaDensity_CGI, fig:Lister_plotBetaDensity_CGI, fig:Seisenberger_plotBetaDensity_CGI, fig:Ziller_plotBetaDensity_CGI} show the kernel density plots of the $\beta$-value distributions stratified by CpG island status for the EPISCOPE, Lister, Seisenberger and Ziller datasets, respectively. These distributions are normalised so that each density has a maximum value of 1.

@@

Because these densities are normalised, these plots don't show the proportion of CpGs in CpG islands. Therefore, Figures \ref{fig:EPISCOPE_plotBetaFreqPoly_CGI, fig:Lister_plotBetaFreqPoly_CGI, fig:Seisenberger_plotBetaFreqPoly_CGI, fig:Ziller_plotBetaFreqPoly_CGI} show the same data but with $\beta$-values grouped into $0.01$-width bins and plotted against the percentage of total CpGs in each bin. These plots highlight that the majority of unmethylated CpGs occur in CpG islands and that most of the intermediate methylation occurs outside of CpG islands, owing to most CpGs being outside of an island.

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Lister data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Ziller_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Ziller data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:EPISCOPE_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Lister data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Lister_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Seisenberger_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Ziller_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Zilelr data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Ziller_plotBetaFreqPoly_CGI}
\end{figure}

@@

CpG $\beta$-values are often computed by aggregating the $m$ and $u$ counts across the forward and reverse strands. On average, this doubles the sequencing coverage of each CpG but presupposes that the two strands are indeed symmetrically methylated. To investigate the validity of this assumption we can compute the correlation of strand-specific $\beta$-values. Tables \ref{tab:EPISCOPE_strand_correlations, tab:Lister_strand_correlations, tab:Seisenberger_strand_correlations, tab:Ziller_strand_correlations} report the Pearson correlation of these strand-specific $\beta$-values for varying sequencing coverage cutoffs.

@@

```{r Ziller_strand_correlations, echo = FALSE, eval = TRUE, message = FALSE, results = "asis"}
x <- read.table("../data/strand_correlation/Ziller_strand_correlations.txt",
stringsAsFactors = FALSE)
colnames(x) <- 1:10
xt <- xtable(x, caption = "Correlations of $\\beta$-values across strands for the Ziller dataset. Each row is a sample and each column is a minimum coverage cutoff.", label = "tab:Ziller_strand_correlations")
align(xt) <- rep("c", ncol(x) + 1)
digits(xt) <- 2
print(xt, booktabs = TRUE, comment = FALSE, tabular.environment = "longtable", floating = FALSE)
```
\subsection{Interpretation}

>>>>

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/bs-seq_protocol.pdf}
\caption{A schematic illustrating several sources of variation in a bisulfite-sequencing experiment. Each sample starts as a population of cells, with potentially different methylation patterns. Each DNA fragment is coloured by its originating cell, although this is unknown in practice. Illustrated are PCR amplication bias (unequal representation of DNA from each cell following PCR amplification), sampling variation (not all DNA fragments are sequenced), alignment errors (not all sequenced fragments are mapped or may be mapped to the wrong location, as is the case for the green fragment), and filtering during methylation calling (not reads are used, the blue read, and reads may have read-positions removed, the black reads).}
\label{fig:bs-seq_protocol}
\end{figure}

@@

\begin{equation*}
Z_{h, i} = \left\{
  \begin{array}{l l}
  (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } \Big (i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th} \Big ) \\
  & \quad \text{ locus on the } h^{th} \text{ fragment}\\
  (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } \Big (i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th} \Big ) \\
  & \quad \text{ locus and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus on the } h^{th} \text{ fragment} \\
  \vdots \\
  (1, 1, \ldots, 1) & \quad \text{if methylated at the } \Big (i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th} \Big ) \\
  & \quad \text{ locus on the } h^{th} \text{ fragment}
  \end{array} \right.
  \end{equation*}

@@

\begin{equation*}
z: z \in \mathcal{R}_{(i, i + 1, \ldots, i + \text{m} - 1)} = \left\{
  \begin{array}{l l}
    (0, 0, \ldots, 0) & \quad \text{if unmethylated at the } \\
    & \quad \Big ( i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th} \Big) \text{ locus}\\
    (0, 0, \ldots, 1) & \quad \text{if unmethylated at the } \\
    \quad \Big (i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 2)^{th} \Big) \text{ locus} \\
      & \quad \text{and methlyated at the } (i + \text{m} - 1)^{th} \text{ locus} \\
    \vdots \\
    (1, 1, \ldots, 1) & \quad \text{if methylated at the } \\
    & \quad \Big (i^{th}, (i + 1)^{th}, \ldots, (i + \text{m} - 1)^{th} \Big ) \text{ locus}
      \end{array} \right.
\end{equation*}

@@

__TODO: Update discussion of results in light of Ziller merged data__

@@

Restricting our attention to CpGs, Figures \ref{fig:EPISCOPE_plotBetaDensity}, \ref{fig:Lister_plotBetaDensity}, \ref{fig:Seisenberger_plotBetaDensity} and \ref{fig:Ziller_merged_plotBetaDensity} show the kernel density estimates of the genome-wide distributions of $\beta$-values, that is, the column-wise summaries, for each sample of the EPISCOPE, Lister, Seisenberer and Ziller datasets, respectively. Figures \ref{fig:EPISCOPE_plotBetaFreqPoly}, \ref{fig:Lister_plotBetaFreqPoly}, \ref{fig:Seisenberger_plotBetaFreqPoly} and \ref{fig:Ziller_merged_plotBetaFreqPoly} show the same data but with the $\beta$-values grouped into $0.01$-width bins and plotted against the percentages of CpGs that fall into each bin.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Lister data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaDensity.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Ziller data. Densities were normalised so that the maximum value for each sample is 1. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included. 'Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_merged_plotBetaDensity}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:EPISCOPE_plotBetaFreqPoly}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Lister data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:Lister_plotBetaFreqPoly}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:Seisenberger_plotBetaFreqPoly}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaFreqPoly.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Ziller data. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Observations have been combined across strands and only CpGs with at least $10\times$ sequencing coverage are included.}
\label{fig:Ziller_merged_plotBetaFreqPoly}
\end{figure}

The bimodality of the genome-wide $\beta$-value distributions is driven by the fact that most CpGs is CpG islands are unmethylated whereas those outside of the islands are mostly methylated. Figures \ref{fig:EPISCOPE_plotBetaDensity_CGI}, \ref{fig:Lister_plotBetaDensity_CGI}, \ref{fig:Seisenberger_plotBetaDensity_CGI} and \ref{fig:Ziller_merged_plotBetaDensity_CGI} show the kernel density plots of the $\beta$-value distributions stratified by CpG island status for the EPISCOPE, Lister, Seisenberger and Ziller datasets, respectively. These distributions are normalised so that each density has a maximum value of 1.

@@

Because these densities are normalised, these plots don't show the proportion of CpGs in CpG islands. Therefore, Figures \ref{fig:EPISCOPE_plotBetaFreqPoly_CGI}, \ref{fig:Lister_plotBetaFreqPoly_CGI}, \ref{fig:Seisenberger_plotBetaFreqPoly_CGI} and \ref{fig:Ziller_merged_plotBetaFreqPoly_CGI} show the same data but with $\beta$-values grouped into $0.01$-width bins and plotted against the percentage of total CpGs in each bin. These plots highlight that the majority of unmethylated CpGs occur in CpG islands and that most of the intermediate methylation occurs outside of CpG islands, owing to most CpGs being outside of an island.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:EPISCOPE_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Lister data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Lister_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Seisenberger_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaDensity_CGI.pdf}
\caption{Kernel density estimates of the genome-wide distribution of CpG $\beta$-values for the Ziller data, stratified by whether the CpG is in a CpG island. Only CpGs with at least $10\times$ sequencing coverage are included. Spikes' in the density estimate are due to the discreteness of $\beta$-values.}
\label{fig:Ziller_merged_plotBetaDensity_CGI}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/EPISCOPE_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the EPISCOPE data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:EPISCOPE_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Lister_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Lister data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Lister_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Seisenberger_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Seisenberger data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Seisenberger_plotBetaFreqPoly_CGI}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{../figures/Ziller_merged_plotBetaFreqPoly_CGI.pdf}
\caption{Frequency polygon of the genome-wide distribution of CpG $\beta$-values for the Zilelr data, stratified by whether the CpG is in a CpG island. $\beta$-values are grouped into $0.01$-width bins and the percentage of CpGs in each bin is plotted on the y-axis. Only CpGs with at least $10\times$ sequencing coverage are included. Percentages are with respect to all CpGs with at least $10 \times$ sequencing coverage, unstratified by CpG island status.}
\label{fig:Ziller_merged_plotBetaFreqPoly_CGI}
\end{figure}

@@

CpG $\beta$-values are often computed by aggregating the $m$ and $u$ counts across the forward and reverse strands. On average, this doubles the sequencing coverage of each CpG but presupposes that the two strands are indeed symmetrically methylated. To investigate the validity of this assumption we can compute the correlation of strand-specific $\beta$-values. Tables \ref{tab:EPISCOPE_strand_correlations}, \ref{tab:Lister_strand_correlations}, \ref{tab:Seisenberger_strand_correlations}, \ref{tab:Ziller_merged_strand_correlations} report the Pearson correlation of these strand-specific $\beta$-values for varying sequencing coverage cutoffs.

@@

```{r Ziller_merged_strand_correlations, echo = FALSE, eval = TRUE, message = FALSE, results = "asis"}
x <- read.table("../data/strand_correlation/Ziller_merged_strand_correlations.txt",
stringsAsFactors = FALSE)
colnames(x) <- 1:10
xt <- xtable(x, caption = "Correlations of $\\beta$-values across strands for the Ziller dataset. Each row is a sample and each column is a minimum coverage cutoff.", label = "tab:Ziller_merged_strand_correlations")
align(xt) <- rep("c", ncol(x) + 1)
digits(xt) <- 2
print(xt, booktabs = TRUE, comment = FALSE, tabular.environment = "longtable", floating = FALSE)
```

\subsection{Interpretation}\label{sec:beta-value_interpretation}

### 2014-12-10 latex/phd_thesis.tex
Minor fixes
Removed Avy chapter from thesis.
Split wgbs_statistical_analysis into wgbs_statistical_framework and wgbs_downstream_analyses
Added tabulary support
Latex fixes

\usepackage{booktabs}

@@

\include{wgbs_statistical_analysis} % wgbs_statistical_analysis.tex

@@

\include{Avy_epialleles} % Avy_epialleles.tex

@@

\chapter{Concluding Comments}

\label{chapter:Conclusions}

% 2 - 4 pages

In this thesis, several fundamental low-level analysis issues for microarrays have been considered.

@@

\label{chapter:AppendixCode}

Insert supplementary figures/R code here.

>>>>

\usepackage{tabulary}

@@

@@

\include{wgbs_statistical_framework} % wgbs_statistical_framework.tex

\include{wgbs_downstream_analyses} % wgbs_downstream_analyses.tex

@@

\chapter{Concluding Comments}\label{chapter:Conclusions}

% 2 - 4 pages
\textbf{TODO}

@@
